Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-02-03 paper/SupMat.Rnw
continued working on the efficient computations section in sup mat

We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the 
beta fecundities lottery model. Defining, for this section, $\beta_i=F_\beta^{-1} \circ \varphi(b_i)$ 
where $F_\beta$ is the CDF of a beta distribution with parameters $0.5,0.5$ and $\varphi$ is the CDF of 
a standard normal distribution, we have $B_i = \eta_i \beta_i$. Then, using the formulas of 
SI section \ref{SIsect:other_mechanisms}, we have
\begin{equation}
E_i = B_i = \eta_i \beta_i 
\end{equation}
\begin{equation}
B_j = \eta_j \beta_j 
\end{equation}
\begin{equation}
C_{i \bs i} = \frac{B_j}{\delta} = \frac{\eta_j}{\delta}\beta_j
\end{equation}
\begin{equation}
\overline{E}_i = \overline{B}_i = \frac{\eta_i}{2} 
\end{equation}
\begin{equation}
\overline{B}_j = \frac{\eta_j}{2} 
\end{equation}
\begin{equation}
\overline{C}_{i \bs i} = \frac{\eta_j}{2 \delta} 
\end{equation}
\begin{equation}
r_{i \bs i} = \ln\left(1-\delta+\delta \frac{B_i}{B_j} \right) = \ln\left(1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}\right)
\end{equation}
\begin{equation}
\epsilon_i^0 = \ln\left( 1-\delta+\delta \frac{\eta_i}{\eta_j}  \right) 
\end{equation}
\begin{equation}
\olep_i^E = \E\left[ \ln\left( 1-\delta+\delta \frac{2\eta_i \beta_i}{\eta_j}  \right)  \right]-\epsilon_i^0
\end{equation}
\begin{equation}
\olep_i^C \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i}{2\eta_j \beta_j}\right)  \right]-\epsilon_i^0
\end{equation}
\begin{equation}
\olep_i^{EC} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]-\olep_i^E-\olep_i^C-\epsilon_i^0
\end{equation}
and
\begin{equation}
\olep_i^{(E\shp C)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right]-\olep_i^E-\olep_i^C-\epsilon_i^0,
\end{equation}
where $(\beta_i^\shp,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_i^\shp),F_\beta^{-1} \circ \varphi(b_j^\shp))$.
Here, $b_i^\shp$ and $b_j^\shp$ are standard normal and independent. Continuing,
\begin{equation}
\olep_i^{(EC)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]
-\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i\beta_i^\shp}{\eta_j \beta_j^\shp} \right) \right].
\end{equation}
Continuing,
\begin{equation}
\olep_i^{[EC]}=\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j} \right) \right]-
\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right],
\end{equation}
where $(\beta_i^\nat,\beta_j^\nat)=(F_\beta^{-1} \circ \varphi (b_i^\nat),F_\beta^{-1} \circ \varphi(b_j^\nat))$,
and $(b_i^\nat,b_j^\nat)$ has the same (standard normal) marginals and covariance as $(b_i,b_j)$ but is
bivariate normal. Note that a choice has been made here: we could have chosen $(b_i^\nat,b_j^\nat)$
to have the same (standard normal) marginals as $(b_i,b_j)$, but with covariance chosen so that
the Pearson, Spearman, or Kendall correlation of $E_i^\nat = \eta_i F_\beta^{-1} \circ \varphi (b_i^\nat)$
and $C_{i \bs i}^\nat = \eta_j F_\beta^{-1} \circ \varphi (b_j^\nat)/\delta$ is the same as that of 
$E_i = \eta_i F_\beta^{-1} \circ \varphi (b_i)$ and $C_{i \bs i} = \eta_j F_\beta^{-1} \circ \varphi (b_j)/\delta$.
But the choice we made seems the most natural under the circumstances.
Continuing, $\olep_i^{[E\nat C]}=\olep_i^{(E C)}-\olep_i^{[E C]}$ is available by subtraction, i.e.,
\begin{equation}
\olep_i^{[E\nat C]} = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right] - 
\E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right].
\end{equation}
This completes the calculation of all the terms of the decomposition 
(\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the beta fecundities lottery model.

We next decompose $\olr_{j \bs i}$ for the beta fecundities lottery model [see (\ref{eq:decomp_overall_rjbsi})]. We have 
\begin{equation}
E_j = B_j = \eta_j \beta_j
\end{equation}
\begin{equation}
C_{j \bs i} = \frac{B_j}{\delta} = \frac{\eta_j \beta_j}{\delta}
\end{equation}
\begin{equation}
\overline{E}_j = \frac{\eta_j}{2}
\end{equation}
\begin{equation}
\overline{C}_{j \bs i} = \frac{\eta_j}{2\delta}
\end{equation}
\begin{equation}
\overline{B}_j = \frac{\eta_j}{2}
\end{equation}
\begin{align}
\olr_{j \bs i} &= \E\left[ \ln\left( 1-\delta+\frac{E_j}{C_{j \bs i}} \right) \right] \\
&= \E\left[ \ln\left( 1-\delta+\frac{B_j}{B_j} \right) \right] \\
&= 0 
\end{align}
\begin{equation}
\epsilon_j^0 = \ln \left( 1-\delta+\delta\frac{\eta_j /2}{\eta_j /2} \right) = 0
\end{equation}
\begin{align}
\olep_j^E &= \E\left[ \ln\left( 1-\delta+\delta\frac{2\eta_j \beta_j}{\eta_j} \right) \right] - \epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+2 \delta \beta_j \right) \right]
\end{align}
\begin{align}
\olep_j^C &= \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_j}{2\eta_j\beta_j} \right) \right]-\epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+\frac{\delta}{2\beta_j} \right) \right]
\end{align}
\begin{align}
\olep_j^{EC} &= \E\left[  \ln\left( 1-\delta+\delta\frac{B_j}{B_j} \right)\right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= -\olep_j^E - \olep_j^C
\end{align}
\begin{align}
\olep_j^{(E\shp C)} &= \E\left[ \ln\left( 1-\delta+\delta\frac{B_j}{B_j^\shp} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right) \right]-\olep_j^E-\olep_j^C,
\end{align}
where $(\beta_j,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_j),F_\beta^{-1} \circ \varphi (b_j^\shp))$ and
$b_j^\shp$ is distributed in the same way as $b_j$ but is independent of it. Continuing,
\begin{align}
\olep_j^{(EC)} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
\E\left[ \ln\left(1-\delta+\delta\frac{B_j}{B_j^\shp}  \right)  \right] \\
&= -\E\left[\ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right)  \right].
\end{align}
Continuing,
\begin{align}
\olep_j^{[EC]} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
\E\left[ \ln\left(1-\delta+\delta\frac{B_j^\nat}{B_j^\nat}  \right)  \right] \\
&= 0.
\end{align}
We showed already (SI section \ref{SIsect:TheoryForLotteryBoth} and 
\ref{SIsect:TheoryForLotteryLogNorm}) that this is $0$ because $E_j$ and $C_{j \bs i}$
are linearly related to each other. Continuing,
\begin{align}
\olep_j^{[E \nat C]} &= \olep_j^{(EC)}-\olep_j^{[EC]} \\
&= -\E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp}  \right)  \right].
\end{align}
This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$
for the beta fecundities lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\eta_i/\eta_j$, but not separately on $\eta_i$ or $\eta_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\eta_i$ or $\eta_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
\begin{equation}
q_{ij} = \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
\end{equation}
For the beta fecundities lottery model, this is
\begin{align}
q_{ij} &= \frac{\eta_i}{\eta_j(1-\delta)+\eta_i \delta} \\
&= \frac{\eta_i/\eta_j}{(1-\delta)+\frac{\eta_i}{\eta_j}\delta}.
\end{align}
Thus the $q_{ij}$ either do not depend on $\eta_i$ or $\eta_j$ at all or they depend on the quotient 
$\eta_i/\eta_j$, rather than depending separately on $\eta_i$ or $\eta_j$. Thus each of the terms of the 
decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\eta_i/\eta_j$ and not separately on $\eta_i$ or $\eta_j$, as stated in the Methods 
section of the main text.

@@

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_1}), as 
$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well

@@

\subsection{Efficient computation for the beta fecundities lottery model}\label{sect:efficient_computation_beta}
%***DAN: fill in later

>>>>

We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the 
beta fecundities lottery model. Defining, for this section, $\beta_i=F_\beta^{-1} \circ \varphi(b_i)$ 
where $F_\beta$ is the CDF of a beta distribution with parameters $0.5,0.5$ and $\varphi$ is the CDF of 
a standard normal distribution, we have $B_i = \eta_i \beta_i$. Then, using the formulas of 
SI section \ref{SIsect:other_mechanisms}, we have
\begin{equation}
E_i = B_i = \eta_i \beta_i 
\end{equation}
\begin{equation}
B_j = \eta_j \beta_j 
\end{equation}
\begin{equation}
C_{i \bs i} = \frac{B_j}{\delta} = \frac{\eta_j}{\delta}\beta_j
\end{equation}
\begin{equation}
\overline{E}_i = \overline{B}_i = \frac{\eta_i}{2} 
\end{equation}
\begin{equation}
\overline{B}_j = \frac{\eta_j}{2} 
\end{equation}
\begin{equation}
\overline{C}_{i \bs i} = \frac{\eta_j}{2 \delta} 
\end{equation}
\begin{equation}
r_{i \bs i} = \ln\left(1-\delta+\delta \frac{B_i}{B_j} \right) = \ln\left(1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}\right)\label{eq:103}
\end{equation}
\begin{equation}
\epsilon_i^0 = \ln\left( 1-\delta+\delta \frac{\eta_i}{\eta_j}  \right) \label{eq:104}
\end{equation}
\begin{equation}
\olep_i^E = \E\left[ \ln\left( 1-\delta+\delta \frac{2\eta_i \beta_i}{\eta_j}  \right)  \right]-\epsilon_i^0 \label{eq:105}
\end{equation}
\begin{equation}
\olep_i^C = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i}{2\eta_j \beta_j}\right)  \right]-\epsilon_i^0 \label{eq:106}
\end{equation}
\begin{equation}
\olep_i^{EC} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]-\olep_i^E-\olep_i^C-\epsilon_i^0
\end{equation}
and
\begin{equation}
\olep_i^{(E\shp C)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right]-\olep_i^E-\olep_i^C-\epsilon_i^0,\label{eq:108}
\end{equation}
where $(\beta_i^\shp,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_i^\shp),F_\beta^{-1} \circ \varphi(b_j^\shp))$.
Here, $b_i^\shp$ and $b_j^\shp$ are standard normal and independent. Continuing,
\begin{equation}
\olep_i^{(EC)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]
-\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i\beta_i^\shp}{\eta_j \beta_j^\shp} \right) \right].
\end{equation}
Continuing,
\begin{equation}
\olep_i^{[EC]}=\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j} \right) \right]-
\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right],\label{eq:110}
\end{equation}
where $(\beta_i^\nat,\beta_j^\nat)=(F_\beta^{-1} \circ \varphi (b_i^\nat),F_\beta^{-1} \circ \varphi(b_j^\nat))$,
and $(b_i^\nat,b_j^\nat)$ has the same (standard normal) marginals and covariance as $(b_i,b_j)$ but is
bivariate normal. Note that a choice has been made here: we could have chosen $(b_i^\nat,b_j^\nat)$
to have the same (standard normal) marginals as $(b_i,b_j)$, but with covariance chosen so that
the Pearson, Spearman, or Kendall correlation of $E_i^\nat = \eta_i F_\beta^{-1} \circ \varphi (b_i^\nat)$
and $C_{i \bs i}^\nat = \eta_j F_\beta^{-1} \circ \varphi (b_j^\nat)/\delta$ is the same as that of 
$E_i = \eta_i F_\beta^{-1} \circ \varphi (b_i)$ and $C_{i \bs i} = \eta_j F_\beta^{-1} \circ \varphi (b_j)/\delta$.
But the choice we made seems the most natural under the circumstances.
Continuing, $\olep_i^{[E\nat C]}=\olep_i^{(E C)}-\olep_i^{[E C]}$ is available by subtraction, i.e.,
\begin{equation}
\olep_i^{[E\nat C]} = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right] - 
\E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right].\label{eq:111}
\end{equation}
This completes the calculation of all the terms of the decomposition 
(\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the beta fecundities lottery model.

We next decompose $\olr_{j \bs i}$ for the beta fecundities lottery model [see (\ref{eq:decomp_overall_rjbsi})]. We have 
\begin{equation}
E_j = B_j = \eta_j \beta_j
\end{equation}
\begin{equation}
C_{j \bs i} = \frac{B_j}{\delta} = \frac{\eta_j \beta_j}{\delta}
\end{equation}
\begin{equation}
\overline{E}_j = \frac{\eta_j}{2}
\end{equation}
\begin{equation}
\overline{C}_{j \bs i} = \frac{\eta_j}{2\delta}
\end{equation}
\begin{equation}
\overline{B}_j = \frac{\eta_j}{2}
\end{equation}
\begin{align}
\olr_{j \bs i} &= \E\left[ \ln\left( 1-\delta+\frac{E_j}{C_{j \bs i}} \right) \right] \\
&= \E\left[ \ln\left( 1-\delta+\frac{B_j}{B_j} \right) \right] \\
&= 0 \label{eq:119}
\end{align}
\begin{equation}
\epsilon_j^0 = \ln \left( 1-\delta+\delta\frac{\eta_j /2}{\eta_j /2} \right) = 0 \label{eq:120}
\end{equation}
\begin{align}
\olep_j^E &= \E\left[ \ln\left( 1-\delta+\delta\frac{2\eta_j \beta_j}{\eta_j} \right) \right] - \epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+2 \delta \beta_j \right) \right] \label{eq:122}
\end{align}
\begin{align}
\olep_j^C &= \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_j}{2\eta_j\beta_j} \right) \right]-\epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+\frac{\delta}{2\beta_j} \right) \right] \label{eq:124}
\end{align}
\begin{align}
\olep_j^{EC} &= \E\left[  \ln\left( 1-\delta+\delta\frac{B_j}{B_j} \right)\right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= -\olep_j^E - \olep_j^C
\end{align}
\begin{align}
\olep_j^{(E\shp C)} &= \E\left[ \ln\left( 1-\delta+\delta\frac{B_j}{B_j^\shp} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right) \right]-\olep_j^E-\olep_j^C, \label{eq:128}
\end{align}
where $(\beta_j,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_j),F_\beta^{-1} \circ \varphi (b_j^\shp))$ and
$b_j^\shp$ is distributed in the same way as $b_j$ but is independent of it. Continuing,
\begin{align}
\olep_j^{(EC)} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
\E\left[ \ln\left(1-\delta+\delta\frac{B_j}{B_j^\shp}  \right)  \right] \\
&= -\E\left[\ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right)  \right].
\end{align}
Continuing,
\begin{align}
\olep_j^{[EC]} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
\E\left[ \ln\left(1-\delta+\delta\frac{B_j^\nat}{B_j^\nat}  \right)  \right] \\
&= 0.\label{eq:132}
\end{align}
We showed already (SI section \ref{SIsect:TheoryForLotteryBoth} and 
\ref{SIsect:TheoryForLotteryLogNorm}) that this is $0$ because $E_j$ and $C_{j \bs i}$
are linearly related to each other. Continuing,
\begin{align}
\olep_j^{[E \nat C]} &= \olep_j^{(EC)}-\olep_j^{[EC]} \\
&= -\E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp}  \right)  \right].\label{eq:134}
\end{align}
This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$
for the beta fecundities lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\eta_i/\eta_j$, but not separately on $\eta_i$ or $\eta_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\eta_i$ or $\eta_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
\begin{equation}
q_{ij} = \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
\end{equation}
For the beta fecundities lottery model, this is
\begin{align}
q_{ij} &= \frac{\eta_i}{\eta_j(1-\delta)+\eta_i \delta} \\
&= \frac{\eta_i/\eta_j}{(1-\delta)+\frac{\eta_i}{\eta_j}\delta}.\label{eq:137}
\end{align}
Thus the $q_{ij}$ either do not depend on $\eta_i$ or $\eta_j$ at all or they depend on the quotient 
$\eta_i/\eta_j$, rather than depending separately on $\eta_i$ or $\eta_j$. Thus each of the terms of the 
decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\eta_i/\eta_j$ and not separately on $\eta_i$ or $\eta_j$, as stated in the Methods 
section of the main text.

@@

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_1}), as 
$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well as at the level of the quantities being estimated.

@@

\subsection{Efficient computation for the beta fecundities lottery model}\label{sect:efficient_computation_beta}

We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the beta fecundities lottery model.

First, for a large integer $M$ (we used $M=\Sexpr{M}$),
carry out the following steps once:
\begin{enumerate}
\item Generate $M$ independent points $(\beta_i^{(k)},\beta_j^{(k)})=(F_\beta^{-1} \circ \varphi (b_i^{(k)}),F_\beta^{-1} \circ \varphi (b_j^{(k)}))$,
$k=1,\ldots,M$, where $(b_i^{(k)},b_j^{(k)})$ are left- or right-tail associated points generated as in SI section \ref{sect:noise},
depending on whether one is currently considering the left- or right-tail associated case.
\item Generate $M$ independent points $(\beta_i^{\shp,(k)},\beta_j^{\shp,(k)})=(F_\beta^{-1} \circ \varphi (b_i^{\shp,(k)}),
F_\beta^{-1} \circ \varphi (b_j^{\shp,(k)}))$, $k=1,\ldots,M$, where $b_i^{\shp,(k)}$ and $b_j^{\shp,(k)}$ are standard normal and 
independent.
\item Generate $M$ independent points 
$(\beta_i^{\nat,(k)},\beta_j^{\nat,(k)})=(F_\beta^{-1} \circ \varphi (b_i^{\nat,(k)}),
F_\beta^{-1} \circ \varphi (b_j^{\nat,(k)}))$, $k=1,\ldots,M$, where the points $(b_i^{\nat,(k)},b_j^{\nat,(k)})$
come from a bivariate normal distribution with standard normal marginals and Pearson correlation the same
as $(b_i^{(k)},b_j^{(k)})$ (see SI section \ref{sect:noise}). 
\end{enumerate}
These data are stored and used repeatedly below.
